---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# auctime

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/auctime)](https://CRAN.R-project.org/package=auctime)
<!-- badges: end -->

The goal of `auctime` is to provide an easy and reproducible way of calculating incremental area under the curve calculations (iAUC) for biomarkers which have measurements taken over at intervals over time. This quantity represents the effect of a treatment on the total concentration of a biomarker.

This is different from the AUC that results from calculation of area under a receiver operating characteristic (ROC) curve, which represents the accuracy of binary classifier.

## Installation

You can install the `auctime` package from GitHub with:

```{r eval=FALSE}
devtools::install_github("scrs-msu/auctime")

library(auctime)
```

## Methods

Several methods are included in calcAUC to find the area under the curve. All methods of calculation use the trapezoidal rule to interpolate between data points.

 - `positive` sums area that is above the baseline (first) measurement, ignoring any area below the baseline (Wolever & Jenkins, 1986)

 - `net` subtracts the area below baseline from the area above baseline (Le Floch et al., 1990)

 - `total` is the area with respect to ground (a baseline of 0)

## Example

Data should be labeled with the biomarker name and timepoint in the format `Biomarker_Time`, with subjects' measurements in rows:

```{r example data}
set.seed(1234)

measurements <- data.frame(Biomarker_0 = rnorm(10,50,20),
                           Biomarker_1 = rnorm(10,70,20),
                           Biomarker_2 = rnorm(10,90,20),
                           Biomarker_3 = rnorm(10,90,20),
                           Biomarker_4 = rnorm(10,70,20),
                           Biomarker_5 = rnorm(10,60,20))

head(measurements)
```

Use `calcAUC()` to calculate area under the curve:

```{r example use, eval=FALSE}
output <- calcAUC(data = measurements)
```

`calcAUC()` produces output containing input data and arguments, subject-wise calculations and plots, a data frame of calculations, and a plot grid of all subjects:

```{r example output, eval=FALSE}
output$input

output$subjects

output$dataframe

output$multiplot
```
